---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
ic <- knitr::all_patterns$md$inline.code
tic <- substr(ic, nchar(ic), nchar(ic))
ntic <- function(n = 4) paste(rep(tic, n), collapse = "")
chlng <- paste0(ntic(4), "{challenge}")
```

# dovetail

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/dovetail)](https://CRAN.R-project.org/package=dovetail)
<!-- badges: end -->

The goal of {dovetail} is to provide a [{knitr}] engine that allows users to write
nested div tags or blockquotes containing executable code as 
[{roxygen2}]-formatted blocks. This will provide a simpler syntax for
construction of [special blockquotes] in Carpentries lessons. These nested
block quotes normally represent solutions that the learner should not see until
they have attempted a given challenge.

# Usage

To use {dovetail}, add `library("dovetail")` in the setup chunk of your Rmarkdown
document and set the engine to one of the following options: `r dovetail:::OUR_TAGS[dovetail:::OUR_TAGS != "end"]`.

# Installation

This project is currently a Work In Progress, but you can install it from the
Carpentries github account:

```r
if (!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
remotes::install_github("carpentries/dovetail")
```

# Example

Consider the [following block quote](https://github.com/swcarpentry/r-novice-gapminder/blob/708ab907306ea764d2e279bb6048c6d9913523c3/_episodes_rmd/06-data-subsetting.Rmd#L263-L282) (taken from [episode 6 of R novice gapminder](http://swcarpentry.github.io/r-novice-gapminder/06-data-subsetting/index.html#challenge-2))
that gives a challenge block with a single solution block nested within it.

Written without the {dovetail} engine, the lesson contributor previously had
to write nested block quotes with kramdown tags like so:

`r invisible(knitr::opts_chunk$set(eval = FALSE))`

````
> ## Challenge 2
> 
> Given the following code:
> 
> ```{r}
> x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
> names(x) <- c('a', 'b', 'c', 'd', 'e')
> print(x)
> ```
> 
> Write a subsetting command to return the values in x that are greater than 4 and less than 7.
> 
> > ## Solution to challenge 2
> > 
> > ```{r}
> > x_subset <- x[x<7 & x>4]
> > print(x_subset)
> > ```
> {: .solution}
{: .challenge}
````

While this solution does provide a way of constructing the nested block quotes
for lessons, it comes with a downside that it's no longer easy to evaluate 
code within these blocks since they are pre-pended by any number of `> ` symbols.

```{r reset, include = FALSE, eval = TRUE}
invisible(knitr::opts_chunk$set(eval = TRUE))
```

With {dovetail}, the above block quote becomes a series of roxygen blocks with
r-code interspersed. The beginning of a nested block is a special `#' @solution`
tag that will indicate the start of a nested block:

``````
`r invisible(knitr::opts_chunk$set(eval = FALSE)); chlng`
#' ## Challenge 2
#' 
#' Given the following code:
#'
#' ```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
#' ```
#'
#' Write a subsetting command to return the values in x that are greater than 4 and less than 7.
#'
#' @solution Solution to challenge 2
#'
#' ```{r}
x_subset <- x[x<7 & x>4]
print(x_subset)
#' ```
```{r tictic, eval = TRUE, results = 'asis', echo = FALSE}
invisible(knitr::opts_chunk$set(eval = TRUE))
cat(ntic(4))
```
``````

The result looks like this:


````markdown
<div class='challenge' markdown='1'>

## Challenge 2

Given the following code:

```r
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
```

```
##   a   b   c   d   e 
## 5.4 6.2 7.1 4.8 7.5
```
Write a subsetting command to return the values in x that are greater than 4 and less than 7.

<div class='solution' markdown='1'>

## Solution to challenge 2
 

```r
x_subset <- x[x<7 & x>4]
print(x_subset)
```

```
##   a   b   d 
## 5.4 6.2 4.8
```

</div>

</div>
````


[{knitr}]:  	https://yihui.org/knitr/
[{roxygen2}]: https://r-lib.github.io/roxygen2
[special blockquotes]: https://carpentries.github.io/lesson-example/04-formatting/index.html#special-blockquotes
