---
title: Test the different knitr engines
---


```{r setup, include = FALSE}
library(dovetail)
library(knitr)
source(dvt_opts())
opts_chunk$set(error = TRUE)
```


```{r some-code}
x  <- "The cat purrrs on my lap."
```


```{callout}
#' ## Count words
#' 
#' 1. How many words are in the sentence `x`?
#'
#' @solution
#'
#' There is more than one way of doing this but here is a solution.
#' 
#' ```{r}
n_words <- length(strsplit(x, " ")[[1]])
n_words
#' ```
#'
#' 1. How many letters are in the sentence `x`?
#' 
#'
#' ```{r}
n_chars <- nchar(gsub("[^a-zA-Z]", "", x))
n_chars
#' ``` 
```

```{challenge}
#' ## Count words
#' 
#' 1. How many words are in the sentence `x`?
#'
#' @solution
#'
#' There is more than one way of doing this but here is a solution.
#' 
#' ```{r}
n_words <- length(strsplit(x, " ")[[1]])
n_words
#' ```
#'
#' 1. How many letters are in the sentence `x`?
#' 
#'
#' ```{r}
n_chars <- nchar(gsub("[^a-zA-Z]", "", x))
n_chars
#' ``` 
```

```{checklist}
#' ## Count words
#' 
#' 1. How many words are in the sentence `x`?
#'
#' @solution
#'
#' There is more than one way of doing this but here is a solution.
#' 
#' ```{r}
n_words <- length(strsplit(x, " ")[[1]])
n_words
#' ```
#'
#' 1. How many letters are in the sentence `x`?
#' 
#'
#' ```{r}
n_chars <- nchar(gsub("[^a-zA-Z]", "", x))
n_chars
#' ``` 
```

```{discussion}
#' ## Count words
#' 
#' 1. How many words are in the sentence `x`?
#'
#' @solution
#'
#' There is more than one way of doing this but here is a solution.
#' 
#' ```{r}
n_words <- length(strsplit(x, " ")[[1]])
n_words
#' ```
#'
#' 1. How many letters are in the sentence `x`?
#' 
#'
#' ```{r}
n_chars <- nchar(gsub("[^a-zA-Z]", "", x))
n_chars
#' ``` 
```

```{keypoints}
#' ## Count words
#' 
#' 1. How many words are in the sentence `x`?
#'
#' @solution
#'
#' There is more than one way of doing this but here is a solution.
#' 
#' ```{r}
n_words <- length(strsplit(x, " ")[[1]])
n_words
#' ```
#'
#' 1. How many letters are in the sentence `x`?
#' 
#'
#' ```{r}
n_chars <- nchar(gsub("[^a-zA-Z]", "", x))
n_chars
#' ``` 
```

```{objectives}
#' ## Count words
#' 
#' 1. How many words are in the sentence `x`?
#'
#' @solution
#'
#' There is more than one way of doing this but here is a solution.
#' 
#' ```{r}
n_words <- length(strsplit(x, " ")[[1]])
n_words
#' ```
#'
#' 1. How many letters are in the sentence `x`?
#' 
#'
#' ```{r}
n_chars <- nchar(gsub("[^a-zA-Z]", "", x))
n_chars
#' ``` 
```

```{prereq}
#' ## Count words
#' 
#' 1. How many words are in the sentence `x`?
#'
#' @solution
#'
#' There is more than one way of doing this but here is a solution.
#' 
#' ```{r}
n_words <- length(strsplit(x, " ")[[1]])
n_words
#' ```
#'
#' 1. How many letters are in the sentence `x`?
#' 
#'
#' ```{r}
n_chars <- nchar(gsub("[^a-zA-Z]", "", x))
n_chars
#' ``` 
```

```{questions}
#' ## Count words
#' 
#' 1. How many words are in the sentence `x`?
#'
#' @solution
#'
#' There is more than one way of doing this but here is a solution.
#' 
#' ```{r}
n_words <- length(strsplit(x, " ")[[1]])
n_words
#' ```
#'
#' 1. How many letters are in the sentence `x`?
#' 
#'
#' ```{r}
n_chars <- nchar(gsub("[^a-zA-Z]", "", x))
n_chars
#' ``` 
```

```{solution}
#' ## Count words
#' 
#' 1. How many words are in the sentence `x`?
#'
#' @solution
#'
#' There is more than one way of doing this but here is a solution.
#' 
#' ```{r}
n_words <- length(strsplit(x, " ")[[1]])
n_words
#' ```
#'
#' 1. How many letters are in the sentence `x`?
#' 
#'
#' ```{r}
n_chars <- nchar(gsub("[^a-zA-Z]", "", x))
n_chars
#' ``` 
```

```{testimonial}
#' ## Count words
#' 
#' 1. How many words are in the sentence `x`?
#'
#' @solution
#'
#' There is more than one way of doing this but here is a solution.
#' 
#' ```{r}
n_words <- length(strsplit(x, " ")[[1]])
n_words
#' ```
#'
#' 1. How many letters are in the sentence `x`?
#' 
#'
#' ```{r}
n_chars <- nchar(gsub("[^a-zA-Z]", "", x))
n_chars
#' ``` 
```


Now that we know that there are `r n_words` words and `r n_chars` characters in the sentence. Let's add these 2 numbers


```{solution}
#' ## add these 2 numbers
#' 
#' @challenge
#' Are these variables available?
#' 
#' ```{r}
#' n_words + n_chars
#' ```
```


```{keypoints}
#' - "Define a function using `name <- function(...args...) {...body...}`."
#' - "Call a function using `name(...values...)`."
#' - "R looks for variables in the current stack frame before looking for them at the top level."
#' - "Use `help(thing)` to view help for something."
#' - "Put comments at the beginning of functions to provide help for that function."
#' - "Annotate your code!"
#' - "Specify default values for arguments when defining a function using `name = value` in the argument list."
#' - "Arguments can be passed by matching based on name, by position, or by omitting them (in which case the default value is used)."
```
