---
title: "Expected Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Expected Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dovetail)
```

This package is designed to be compatible with existing workflows, meaning that
one should be able to mix the current syntax with new syntax seamlessly:

# Example 1: nested block quote

<div class="row" markdown="1">

<div class="col-md-6" markdown="1">

## With New Formatting

```{challenge, coo = TRUE}
#' ## Challenge 2
#' 
#' Given the following code:
#'
#' ```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
#' ```
#'
#' Write a subsetting command to return the values in x that are greater than 4 and less than 7.
#'
#' @solution Solution to challenge 2
#'
#' ```{r}
x_subset <- x[x<7 & x>4]
print(x_subset)
#' ```
```

</div>
<div class="col-md-6" markdown="1">

## With Old Formatting

````
> ## Challenge 2
> 
> Given the following code:
> 
> ```{r}
> x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
> names(x) <- c('a', 'b', 'c', 'd', 'e')
> print(x)
> ```
> 
> Write a subsetting command to return the values in x that are greater than 4 and less than 7.
> 
> > ## Solution to challenge 2
> > 
> > ```{r}
> > x_subset <- x[x<7 & x>4]
> > print(x_subset)
> > ```
> {: .solution}
{: .challenge}
````

> ## Challenge 2
> 
> Given the following code:
> 
> ```{r}
> x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
> names(x) <- c('a', 'b', 'c', 'd', 'e')
> print(x)
> ```
> 
> Write a subsetting command to return the values in x that are greater than 4 and less than 7.
> 
> > ## Solution to challenge 2
> > 
> > ```{r}
> > x_subset <- x[x<7 & x>4]
> > print(x_subset)
> > ```
> {: .solution}
{: .challenge}


</div>
</div>


# Options

It's possible to set different options for each chunk just like you would any
other Rmarkdown chunk. For example, this following chunk is provisioned with:
 `{callout, eval = FALSE}`


```{callout, eval = FALSE, coo = TRUE}
#' The code in this callout chunk will not be evaluated, but the text will
#' still appear!
#' ```{r}
stop("hammertime")
#' ```
```
<!-- NB: I have to cheat a little here -->
```{callout, eval = FALSE}
#' The code in this callout chunk will not be evaluated, but the text will
#' still appear!
#' ```{r}
stop("hammertime")
#' ```
```


## Overriding options

You can override options within the chunk. This following chunk is provisioned
with `{discussion, echo = FALSE}`

```{discussion, coo = TRUE, echo = FALSE}
#' All chunks within the dovetail chunk **will be formatted as markdown**
#' so you can write your lessons with ease.
#' 
#' Moreover:
#' 
#'
#' - chunk options are inherited
#' - objects within chunks persist in the environment
#' - you can modify chunk options within R chunks
#'
#' 
#' ### Example: a sentence
#' 
#' ```{r}
cat("THERE IS A HIDDEN CHUNK HERE!")
x <- "the cat purrrs on my lap."
#' ```
#'
#' ```{r echo = TRUE}
n <- length(strsplit(x, " ")[[1]]) # how many words does "x" have
n
#' ```
#'
#' A sentence that has **`r n`** words is "`r x`"
```

The sentence "`r x`" has **`r n`** words.

## Jekyll options

The Carpentries lessons currently run on Jekyll, so we have provided hooks to
modify the output to work with jekyll, which can be loaded with the `dvt_opts()`
function:

```{r}
source(dvt_opts())
```

```{challenge}
#' here is some R code
#' ```{r}
1 + 1
#' ```
```

```{r echo = FALSE}
  knitr::opts_knit$set(knitr::opts_knit$get(default = TRUE))
  knitr::opts_chunk$set(knitr::opts_chunk$get(default = TRUE))
  knitr::knit_hooks$set(knitr::knit_hooks$get(default = TRUE))
```

The code for this is migrated from the styles repository:

````r
```{r echo = FALSE, comment = NA, results = 'asis'}
cat(readLines(dvt_opts()), sep = "\n")
```
````


# Preparing an existing lesson for using {dovetail}:

To update the lessons, the special blockquotes will be converted to code chunks and the setup chunk will do the following replacement:

```diff
-source('../bin/chunk-options.R')
+library('dovetail')
+source(dvt_opts())
```

To prepare an existing lesson with RMarkdown, install {dovetail} and {pegboard}
with

```r
remotes::install_github("carpentries/pegboard", dep = TRUE)
remotes::install_github("carpentries/dovetail", dep = TRUE)
```

> **A note on {pegboard}**: This package will read in and transform your
> existing lessons to the dovetail format, but there may be rendering mistakes, 
> so please inspect the result with git diff and rendering the site locally

There are two steps to updating the lessons:

 1. open your lesson in RStudio
 2. run `source(system.file("convert", "convert.R", package = "dovetail"))`
 
The conversion script looks like this:

````r
```{r echo = FALSE, comment = NA, results = 'asis'}
cat(readLines(system.file("convert", "convert.R", package = "dovetail")), sep = "\n")
```
````
